Reproducible Research: Peer Assessment 1 (bbakerharvey)
=======================================================

## Loading and preprocessing the data
First we have to load the data.  Since it is presented as a 
Zip (archive) file format, it must be unzipped, then loaded
into the working data.frame in R.  We plan to use data.table
funcitons, so convert to data.table format, too.

```{r}
library("data.table")
unzip("activity.zip")
stepdata <- data.table(read.table("activity.csv",sep=",",
                                  header=TRUE))
str(stepdata)
```
The raw data has been loaded into the data.frame stepdata.

Since we know we have NA values in the dataset, first let's look
at exploring the data with the NA values removed.

```{r}
stripna=stepdata[!is.na(stepdata$steps),]
str(stripna)
```
## What is daily total number of steps taken?

Using the abbreviated dataset, lets look at the total steps
reported for each day.  In table form this is:

```{r}
library("plyr")
withdaily <- ddply(stripna,~date,summarise, dailysteps=sum(steps))
str(withdaily)
withdaily
```

Let's view this as a histogram:
```{r}
hist(withdaily$dailysteps,xlab="steps",
     main="Distribution of Daily Step Counts", col="steelblue")
```

Most days seem to be 10,000 to 15,000 steps.

## What is mean total number of steps taken per day?

To see more about daily steps, let's compute mean and median
of the daily step counts

```{r}
dailymean=mean(withdaily$dailysteps)
dailymedian=median(withdaily$dailysteps)

cat(sprintf("mean number of steps is %f and median is %f\n",
            dailymean,dailymedian))

```


## What is the average daily activity pattern?

Consider the average activity pattern for different times during the day.  

```{r}
dailypattern <- ddply(stripna,~interval,summarise, 
                      periodsteps=mean(steps))
str(dailypattern)
```
So let's look at this as a time series plot:

```{r}
plot(dailypattern$interval,dailypattern$periodsteps,
     xlab="Interval",ylab="Average Steps",
     type="l")
```

In the graph, there appears to be a peak value.

```{r}
highest = max(dailypattern$periodsteps)
which = dailypattern$interval[dailypattern$periodsteps==highest]

cat(sprintf("peak value of %f is found in interval %5.0f\n",
            highest,which))
```


## Imputing missing values

We know our data set had some missing values, which we have
omitted from the analysis until now.  Now let's consider
filling in those blank observations as we consider looking
at weekly patterns.  For the analysis we will use a simple
strategy of replacing the NA with an average number of
steps for the missing interval.  Call this processed step count
procsteps.

```{r}
avgvals<- dailypattern$periodsteps
str(avgvals)
avgvals
stepdata$isna <- is.na(stepdata$steps)
stepdata$procsteps <- 0.0 + stepdata$steps
loopmax <- length(stepdata$date)
print(loopmax)
for (i in 1:loopmax) {
        if (stepdata[i]$isna) {
                stepdata[i]$procsteps <- 
                        dailymean / ( 2400.0 / 5.0)
                }
        }
str(stepdata)

```

### For this new imputed dataset, consider the daily activity
```{r}
library("plyr")
procdaily <- ddply(stepdata,~date,summarise,
                   dailyprocsteps=sum(procsteps))
str(procdaily)
procdaily
```

To see more about daily steps, let's compute mean and median
of the imputed daily step counts

```{r}
procmean=mean(procdaily$dailyprocsteps)
procmedian=median(procdaily$dailyprocsteps)

cat(sprintf("For this proc dataset: mean number of steps is %f and median is %f\n",
            procmean,procmedian))

```

This can be seen as a histogram:
```{r}
hist(procdaily$dailyprocsteps,xlab="procsteps",
     main="Distribution of Daily Step Counts (imputed)", 
     col="pink")
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
vsat <- weekdays(as.Date(stepdata$date)) == "Saturday"
vsun <- weekdays(as.Date(stepdata$date)) == "Sunday"
stepdata$iswe <- vsat | vsun
sum(stepdata$iswe)
```
Compute the weekend and weekday patterns
```{r}
weekendpattern <- ddply(stepdata[stepdata$iswe==TRUE],~interval
        ,summarise, periodsteps=mean(procsteps))
str(weekendpattern)

weekdaypattern <- ddply(stepdata[stepdata$iswe==FALSE],~interval
        ,summarise, periodsteps=mean(procsteps))
str(weekdaypattern)
```
Plot the weekend and weekday patterns
```{r}
par(mfcol=c(2,1))

plot(weekendpattern$interval,weekendpattern$periodsteps,
     xlab="Interval",ylab="Average Steps", main="Weekend",
     type="l")

plot(weekdaypattern$interval,weekdaypattern$periodsteps,
     xlab="Interval",ylab="Average Steps", main="Weekday",
     type="l")

```   
     
