Reproducible Research: Peer Assessment 1 (bbakerharvey)
=======================================================

## Loading and preprocessing the data
First we have to load the data.  Since it is presented as a 
Zip (archive) file format, it must be unzipped, then loaded
into the working data.frame in R.  We plan to use data.table
funcitons, so convert to data.table format, too.

```{r}
library("data.table")
unzip("activity.zip")
stepdata <- data.table(read.table("activity.csv",sep=",",header=TRUE))
str(stepdata)
```
The raw data has been loaded into the data.frame stepdata.

Since we know we have NA values in the dataset, first let's look
at exploring the data with the NA values removed.

```{r}
stripna=stepdata[!is.na(stepdata$steps),]
str(stripna)
```
## What is daily total number of steps taken?

Using the abbreviated dataset, lets look at the total steps
reported for each day.  In table form this is:

```{r}
library("plyr")
withdaily <- ddply(stripna,~date,summarise, dailysteps=sum(steps))
str(withdaily)
withdaily
```

Let's view this as a histogram:
```{r}
hist(withdaily$dailysteps,xlab="steps",
     main="Distribution of Daily Step Counts", col="steelblue")
```

Most days seem to be 10,000 to 15,000 steps.

## What is mean total number of steps taken per day?

To see more about daily steps, let's compute mean and median
of the daily step counts

```{r}
dailymean=mean(withdaily$dailysteps)
dailymedian=median(withdaily$dailysteps)

cat(sprintf("mean number of steps is %f and median is %f\n",
            dailymean,dailymedian))

```

This is an interesting result because each day includes many 
periods with zero activity, that is no steps taken, so the 
median (middle observation) is in fact zero for all days.

Let's just see how many zero activity observations there are
each day.

```{r}
ddply(stripna,~date,summarise, meansteps=mean(steps,na.rm=TRUE),
      mediansteps=median(steps),obscount=length(steps),
      zerocount=sum(steps==0))
```


## What is the average daily activity pattern?

Consider the average activity pattern for different times during the day.  

```{r}
dailypattern <- ddply(stripna,~interval,summarise, 
                      periodsteps=mean(steps))
str(dailypattern)
```
So let's look at this as a time series plot:

```{r}
plot(dailypattern$interval,dailypattern$periodsteps,
     xlab="Interval",ylab="Average Steps",
     type="l")
```

In the graph, there appears to be a peak value.

```{r}
highest = max(dailypattern$periodsteps)
which = dailypattern$interval[dailypattern$periodsteps==highest]

cat(sprintf("peak value of %f is found in interval %5.0f\n",
            highest,which))
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
